The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-09-20 adaptation/article/article.tex
Update on Overleaf.

\title{The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery}
\shorttitle{Context-inference-dependent motor learning}
\author[1]{Cuevas Rivera, Darío}
\author[1]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
\affiliation{~}

@@

\abstract{Abstract}
Humans are able to adapt motor commands to changing dynamics of the environment and of the body itself. While we must adapt to slow changes in our body dynamics, e.g. because of muscle exhaustion throughout the day, we can also adapt to sudden changes in the environment. For example, when picking up a heavy objects, we must adapt our motor commands to be able to move the object around. The size and material of an object provide us with contextual information which we can use to prepare the correct response to moving the object, and our ability to switch between clear contexts (e.g. between lifting a big rock and small piece of paper) has been well studied. However, contextual information is not always clear or reliable (e.g. a big object of unknown material). In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection. We show how our model explains the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.

\section{Introduction}
It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}. Additionally, humans have been shown to dynamically switch between different learned tasks. What prompts these switches is still a matter of research.

In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement. For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error. Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.

@@

\section{Results}
In this work, we present a motor adaptation model in which the agent (e.g. a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation. Several models exist for error-based motor adaptation, but we expand on the existing models by adding an explicit component of context inference which guides the selection of the adequate internal model, its updating and the sampling of actions.

We used our model to explain several experimental phenomena found in previous studies. We simulated representative experiments from these studies to show how our model explains their findings using the dynamics of context inference. In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss how context inference explains the data.

Before presenting our simulations, we will briefly explain the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.

\subsection{Modeling context-dependant adaptation}
The model we present has three main components: (1) context-inference, (2) motor adaptation and (3) action selection. Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference. A schematic representation of the model can be seen in \fref{fig:model}.

\begin{figure}
\centering
\includegraphics[]{./figures/figure_1.png}
\caption{Description of the model. (A) Definition of a context in terms of its components. (B) Update equations for context inference, motor adaptation and action selection. $\zeta$ represents the identity of the context as a categorical variable, $\vartheta$ are observations, $c_t$ are motor commands at trial $t$, $\phi$ are the parameters of the forward model and $s$ are the hidden states of the environment (e.g. the position of the hand). Note that because action selection is not performed via Bayes' theorem, the notation $p(c_t)$ is used. (C) Schematic representation of the model. At trial $t$, an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$). After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process repeated. Note that the posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
\label{fig:model}
\end{figure}

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. To see how these three components affect each other, see \fref{fig:model}B and C.

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors. For example, we can learn to perform a reaching movement in front of us underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. We chose to make use of an exact Bayesian learner. As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, we make use of the example of reaching movements. Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \phi) \label{eqn:forward-model}
\end{equation}
where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the hand during the reaching movement), $s$ is the current state (e.g. the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$. Throughout our lives, the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.

When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that:
\begin{equation}
q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi)
\end{equation}
where $q(\phi | ...)$ represents the new estimate of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$. The likelihood $p(\vartheta | s, c, \phi)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.

\subsection{Context inference}
In our model, the idea of context is quite general (see \fref{fig:model}A). The context comprises the relevant elements of the environment (e.g. the strength and direction of the wind when walking), of the task at hand (immediate and long-term goals, reward structure, etc) and task-related information internal to the decision-making agent (e.g. forward model, decision-making algorithms, motor adaptation equations, expected observations). Previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.

Identifying any possible context with a categorical variable $\zeta$, the identification of a context is done by our model through Bayesian inference:
\begin{equation}
q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
\end{equation}
where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall estimation of observing any one context, if the context did change. Below, we discuss how these two components effect different experimentally observed phenomena.

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions (e.g. the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them. For this section, the most important component of $\vartheta$ is contextual sensory information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the contextual sensory information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.

By making use of NormalGamma priors on context, the model performs exact Bayesian inference at every trial. The exact update equations can be found in the Methods section.

\subsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model. In other words, a distribution is created as follows:
\begin{equation}
p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta) \label{eq:action-selection}
\end{equation}
where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.

\section{Experimental results}
In this section, we go through a number of experimentally-observed phenomena and show that context inference, as done in our model, provides a unifying explanation for all of them.

For clarity, we first introduce necessary terminology. As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force. Depending on the trial, the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise. Additionally, abusing notation, we can say that $B = (-A)$, as the forces have the same magnitude but point in opposite directions. We can also talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g. clockwise), but the second one has half the magnitude. Finally, error-clamp trials, in which the mechanical arm forces the participant to make straight-line movements, are represented with the letter E.

@@

\subsection{Savings and slow/fast switching}
Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

In this section, we discuss savings, and the related phenomenon of deadaptation, in terms of the switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019}, but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.

To show this, we categorized experiments by looking at the amount of contextual information made available to participants. In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants \citep[e.g.][]{Davidson_Scaling_2004,Zarahn_Explaining_2008} in the form of large prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.

To our knowledge, there are no experiments in which no contextual information is available. However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information. These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is still available to participants. In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.

To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies which differ in the amount of contextual information. \cite{Kim_Neural_2015} performed a visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results is quite evident, both in the immediacy of the switches as well as in the fact that learning continues from the state it was when the context was last observed.

\begin{figure}
\centering
\includegraphics{./figures/figure_2.png}
\caption{Savings and de-adaptation. Data from our simulations (left column) compared to data from \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} (right column). In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number. The black lines represent the optimal adaptation, i.e. the size of the visuomotor rotation during the task. (A) Experiment by \cite{Kim_Neural_2015} with two rotations (-40 and 40 degrees), in addition to baseline. The switches between contexts were visually cued and therefore switches are immediate, as can be seen in both simulations (left) and the experimental data from \cite{Kim_Neural_2015} (right). (B) An experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements. It can be seen in both the simulations (left) and experimental data from \cite{Oh_Minimizing_2019} (right) that after context switches (e.g. at trial 120), the agent returns to the adaptation of the previous block, with one or two trials of delay. (C) Same experiment as (B) but  with a 10 degree adaptation. In this case, savings are much slower.} \label{fig:oh-2019}
\end{figure}

\cite{Oh_Minimizing_2019} performed a similar experiment in which contexts were not visually cued. In addition, the authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model. Participants in \cite{Oh_Minimizing_2019} showed immediate switching in the first experiment, while switching was much slower in the second experiment. As can be seen on the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low. This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly identifying the context from the noise due to trial-to-trial variation in motor output. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

@@

\section{Slower adaptation in volatile and unpredictable environments}
In our model, adaptation is gated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.

>>>>

\title{The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery}
\shorttitle{Context-inference-dependent motor learning}
\author[1,2]{Cuevas Rivera, Darío}
\author[1,2]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
\affiliation{~}

@@

\abstract{Abstract}
 Humans are able to adapt their movements to changes in the dynamics of the environment and of their own bodies, a phenomenon known as motor adaptation. Motor adaptation has been shown to be contextual in experiments in which participants must learn to adapt to different altered dynamics, and must be able to switch between the learned adaptations dynamically. In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection. We show how our model provides a unifying explanation for the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.

\section{Introduction}
Humans are able to adapt movements to changing dynamics of the environment and of the body itself. While we must adapt to slow changes in our body dynamics, e.g. because of muscle exhaustion throughout the day, we can also adapt to sudden changes in the environment. For example, when picking up a heavy object, we must adapt our motor commands to be able to move the object around. Even before picking it up, the size and material of an object provide us with contextual information which we can use to prepare the correct posture for moving the object. However, contextual information cannot always be relied on, e.g. a big object of unknown material.

It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}, which slows down learning. Additionally, humans have been shown to dynamically switch between different learned tasks \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}. The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the way in which switching can be impeded or improved, are still an active topic of research.

In order to study motor adaptation, motor error is defined in experiments as the deviation between the observed movements of a participant and a baseline movement. For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error. Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.

@@

\section{Results}
In this work, we present a motor adaptation model in which the agent (e.g. a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation. Several models exist for such error-based motor adaptation, but we expand on existing models by adding an explicit component of context inference. As we will show, this context inference guides the selection of the adequate internal model, its updating and the sampling of actions.

In what follows, we show that the proposed model explains several experimental phenomena found in previous studies. We simulated representative experiments from these studies to illustrate how our model explains these experimental findings using the dynamics of context inference. In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss in detail how context inference explains the experimental phenomena.

Before presenting our simulations, we will briefly go through the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.

\subsection{Modeling context-dependent adaptation}
The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these component happen in this order, and each component informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.

Central to the model is the concept of context. In our model, contexts are defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the variables of the environment. Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to identify the context. For example, when lifting an object of unknown weight, an agent might have learned one context for heavy objects and one for light objects. When observing an object to be lifted, the agent can use its size and texture to estimate the weight of the object, which in turn allows the agent to infer the appropriate context and, with it, decide how to lift the object. These contextual cues (size and texture) do not necessarily identify the context (e.g. it could be a movie prop that looks like a rock).

\begin{figure}
\centering
\includegraphics[]{./figures/figure_1.png}
\caption{Description of the model. (A) Definition of a context in terms of its components. (B) Update equations for context inference, motor adaptation and action selection. $\zeta$ represents the identity of the context as a categorical variable, $\vartheta$ are observations, $c_t$ are motor commands at trial $t$, $\phi$ are the parameters of the forward model and $s$ are the hidden states of the environment (e.g. the position of the hand). (C) Schematic representation of the model. The arrows show flow of information, i.e. the dependencies in the update equations, as well as the flow of a single trial. The elements (e.g. $q(\zeta)$) are those shown in B. Trial $t$ starts at the top when an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$). After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process is repeated. The posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
\label{fig:model}
\end{figure}

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. For an overview, \fref{fig:model}B and C shows how these three components affect each other.

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors. For example, we can learn to perform a reaching movement even when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. Here, we use this principle and chose to make use of an exact Bayesian learner. As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials. As we will show below this feature will be useful to explain specific key phenomena parsimoniously. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, we make use of the example of reaching movements. It is now widely assumed that humans learn associations between motor commands and outcomes in different environments \cite{Wolpert_Multiple_1998}. These associations, called forward models, are used by humans for action selection. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \phi) \label{eqn:forward-model}
\end{equation}
where $\vartheta$ are the outcomes of an action (e.g. the final position of the hand in a reaching movement), $s$ is the current state (the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$. We assume that during learning the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.

When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that
\begin{equation}
q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi)
\end{equation}
where $q(\phi | ...)$ represents the new estimated probability distribution over the possible values of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$. The likelihood $p(\vartheta | s, c, \phi)$ is the probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.

\subsubsection{Context inference}
In our model, the idea of context is quite general (see \fref{fig:model}A). The context comprises the relevant elements of the environment (e.g. the strength and direction of the wind when walking), of the task at hand (e.g. immediate and long-term goals and reward structure) and task-related information internal to the decision-making agent (e.g. forward model, decision-making algorithms, motor adaptation equations, expected observations). Most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context. In addition, \cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.

Following the these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.

Context is a categorical variable $\zeta$, where the identification of a context is done through Bayesian inference:
\begin{equation}
q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
\end{equation}
where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall expectation of observing any one context, if the context did change. Below, we discuss how these two sources of information effect different experimentally observed phenomena.

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions (e.g. the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them, as is the case for visual cues in experiments \cite[e.g.][]{Addou_Colored_2011}. For this section, the most important component of $\vartheta$ is contextual sensory information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to infer the identity of a context.

\subsubsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model. In other words, a distribution is created as follows:
\begin{equation}
p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta) \label{eq:action-selection}
\end{equation}
where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.
\todo{Maybe add a section with the progression of a trial}
\section{Experimental results}
In this section, we present experimentally-observed phenomena in four sections, and show that context inference, as done in our model, provides a unifying explanation for all of them. In the first section, we discuss switches between contexts, and how slow context inference affects these switches. In the second section, we focus on the effects of volatility in the environment on motor adaptation. In the third section, we discuss the effects of slow context inference on action selection. Finally, in the fourth section we discuss context inference during error-clamp trials, and its effect on behavior.

For clarity, we first introduce necessary terminology that is typically used in experimental studies. As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force (i.e. a force that pushes the participant's hand in the direction perpendicular to movement). Depending on the trial, the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise. Additionally, abusing notation, we can say that $B = (-A)$, as the forces have the same magnitude but point in opposite directions. Studies often talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g. clockwise), but the second one has half the magnitude. Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter E.

@@

Additionally, we use the term "learned adaptation" and similar to refer to motor outputs (e.g. reaching movements) that have already been adapted by the participant to a change in the dynamics of the environment. For example, the learned adaptation during a task with a curl force refers to the movements of the participant that already counteract the force exerted by the mechanical arm.

\subsection{Savings and slow/fast switching}
The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019} but we show that expanding forward model switching into context inference explains facets of savings that these previous models do not.

To show this, we examined multiple experimental studies in which savings is displayed. We categorized these studies based on the amount of contextual information made available to participants. In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants \citep[e.g.][]{Davidson_Scaling_2004,Zarahn_Explaining_2008} in the form of large prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.

To our knowledge, there are no experiments in which no contextual information is available. However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information. These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is nevertheless available to participants. In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.

To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \cite{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information. \cite{Kim_Neural_2015} performed a visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e. when the solid black line (representing the real context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches. However, adaptation is incomplete (i.e. the blue line never reaches 40 or -40) in both experimental data and simulations until late in the experiment (around trial 500).

\begin{figure}
\centering
\includegraphics{./figures/figure_2.png}
\caption{Savings and de-adaptation. Data from our simulations (left column) compared to data from \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} (right column). Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown. In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number. The black lines represent the optimal adaptation, i.e. the size of the visuomotor rotation during the task. (A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O. Participants must adapt to the rotations during shooting movements. The y-axis is in degrees; the black line represents the real visuomotor rotation, at either zero or 20 degrees, and the blue line shows the participants' adaptation (averaged across all participants), which can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation. When black and blue lines match, the participant perfectly counteracted the visuomotor rotation and reached the target. (B) An experiment by \cite{Oh_Minimizing_2019} similar to A, but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced. O-A and A-O transitions (i.e. adaptation and de-adaptation) can be observed. Blue and black lines are as in A. (C) Same experiment as (B) but with a 10 degree adaptation.} \label{fig:oh-2019}
\end{figure}

\cite{Oh_Minimizing_2019} performed partially-cued experiment similar to \cite{Kim_Neural_2015}. The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model. Participants in \cite{Oh_Minimizing_2019} showed immediate switching in the first experiment, which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (real rotation). In the second experiment, switching happens more slowly, with adaptation (blue line) lagging behind the switches in real context (black line), and slowly catching up. As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low. This difference is explained by our model in terms of the size of the adaptation: as the size of the adaptation is lower (10 degrees), it becomes more difficult to distinguish errors made by incorrectly identifying the context from the noise due to trial-to-trial variation in motor output. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

@@

\subsection{Slower adaptation in volatile and unpredictable environments}
One central element in our model is that adaptation is gated by the uncertainty on the current context: a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.
